Fusion Moves for Correlation Clustering

### 2014-11-12 paper/egpaper_for_review.tex
exp

@@

\subsection{Fast Optimization of CCFusion Moves} \label{sec:fast_optimization}
In general the auxiliary fusion problem~\ref{eq:fusion_move_a} is still NP-hard.
However, many variables have been fixed to be zero and we can reformulate~\ref{eq:fusion_move_a}
into a correlation clustering problem on a coarser graph, where all nodes which are connected
via must-link constraints are merged into single nodes. We call this graph \emph{contracted graph}.
%
\begin{definition}\emph{(Contracted Graph)}
Given a weighted graph $G=V,E,w)$ segmentation of $G$ given by $y\in P(G)$,
we denote by $S: V \to \mathbb{N}$ a mapping from the nodes to the segments.
We define the contraction of graph $G_y=(V_y,E_y, \bar{w})$
by $V_y=\{S(v)|v\in V\}$, $E_y=\{S(u)S(v)|uv\in E\}$, and 
$\forall \bar{u}\bar{v}\in E_y: \bar{w}_{\bar{u}\bar{v}}=\sum_{uv\in E, S(u)=\bar{u},S(v)=\bar{v}}w_{uv}$ 
\end{definition}
%
Any clustering $\bar{y}$ of the contracted graph $G_y=(V_y,E_y)$ can be \emph{back projected} to a clustering $y$ the original graph $G=(V,E)$ by
%
\begin{align}
y_{uv}&=\left\{ 
\begin{array}{ll}
\bar{y}_{S(u)S(v)}& \textrm{ if }S(u) \neq S(v)\\
0                & \textrm{ else}
\end{array}
\right.& \forall uv \in E
\end{align}
Furthermore, the back projection of the optimal segmentation $\bar{y}$ of the contracted graph is a solution of problem~\ref{eq:fusion_move_a}.

@@

As discussed in~\cite{Lempitsky-2010}, proposals
should have two properties: \emph{quality} 
and \emph{diversity}.
%
A proposal have a high quality if it at least loacly
The quality of the individual proposals
ensures low energy solutions since the
result of the fusion cannot be higher then
the energy of the current best solution.
%
Diversity between the individual proposals
increase the chances to escape from local minima.
%
For correlation clustering fusion we add a third
property: \emph{size}.
The size of the contraction graph directly
depends on the number of connected components
in the proposal solution and the current best solution.
The extreme  case is a proposal where each 
node is in a separate connnected component.
Solving the fusion move with such a proposal
is equivalent to solving the original problem.
Solving the fusion move with a proposal with 
a single connected component will result 
in the current best solution.
Therefore the size of the proposals
should be small enough, such that solving
eq.~\ref{eq:edgeproblem} can
be done fast enough, but large enough to make
meaningful moves (??? this last sentence needs to be rewritten).

@@

\noindent \textbf{Randomized HierarchicalClustering (RHC):}
%\subsubsection{Randomized HierarchicalClustering (RHC)}
%
To generate fast energy aware proposals we can use energy
based hierarchical clustering (EHC) as defined in sec.~\ref{sec:ehc}.
EHC follows the energy function, therefore the \emph{quality}
of the proposals is high.
To get \emph{diversity} among the different proposal,
we add normal distributed noise  $\mathcal{N}(0, \sigma_{ehc})$
to each edge weight.
To get proposals of the desired \emph{size}, we use a
different stop condition in EHC, and stop iff a certain number of connected
components is reached.

@@

\noindent \textbf{Randomized Watersheds (RWS):}
%\subsubsection{Randomized Watersheds (RWS)}
Watersheds have become quite popular for
graph segmentation and have a strong connection
to energy minimization~\cite{couprie_2011}.
%
The edge weighted watershed algorithm~\cite{meyer_2013}
with random seeds can be used to get
cheap proposals.
To improve \emph{quality} we do not use $n$ seeds distributed uniformly
over all nodes but use the following.
%
Draw $n/k$ negative edges,  and give the endpoints a different seed.
Doing so, a random subset of negative edges is forced
to be cut within each proposal.
For additional \emph{diversity}, noise $\mathcal{N} (0, sigma_{ws})$
can be added to
the edge weighs.

@@

%-----------------------------------------------------------------------
\subsection{Limitations}

@@

For \textbf{CGC}~\cite{beier_2014_cvpr} we used the public available implementation\footnote{\label{opengm-cvpr2014}
\small{{https://github.com/opengm/opengm/tree/cgc-cvpr2014}}
}
of the Beier~\etal and for \textbf{KL}~\cite{Kernighan-1970} the implementation in OpenGM\footlabel{opengm-master}{\url{https://github.com/opengm/opengm}}.
For integer multicuts (\textbf{MC-I}) and relaxed multicuts (\textbf{MC-R})~\cite{kappes_2013_arxiv} we try the public available implementation in OpenGM\footref{opengm-master}
of Kappes~\etal. We realize that their implementation has some limitations on large problems and make the following modifications.
Firstly, we use \emph{index-min-heap}~\cite{Sedgewick-2011} within the shortest path search by the Dykstra algorithm - this gives a speed up for MC-R.
Secondly, we follow~\cite{kroeger_2012_eccv} and search for shortest paths and add those only if they are non-chordal instead of searching for the shortest non-chordal path
during the separation procedure. Fig.~\ref{fig:at:knott-450b} show the improvements with our modifications compared to the implementation in OpenGM for the knott-3d-450 dataset.
This procedure is one magnitude faster, but might cause a few extra outer iterations.
In~\cite{kroeger_2012_eccv}. this has been used for MC-I only. 
For MC-R this search procedure is not sufficient and need to be followed by a search for a shortest non-chordal paths.
From the field of data-mining we compare to the PIVOT-algorithm~\cite{Ailon-2008} followed by a round of BOEM~\cite{Gionis-2007} denoted by \textbf{PIVOT-BOEM}.
We use the public available code of Micha Elsner\footnote{\url{http://www.ling.ohio-state.edu/~melsner/resources/correlation-readme.html}}.
While this implementation uses full adjacency matrixes it does not scale as well as the algorithm can and therefor can not be applied on all datasets.
We also run classical fusion moves~\cite{Lempitsky-2010} (\textbf{Fusion}) and select distinct labels for the two candidate segmentations.

@@

We compare this methods to the following methods suggested in the present paper.
%
Energy Based Based Hierarchical Clustering (\textbf{EHC})
as described in sec.~\ref{sec:ehc}.
%
CGC warm started with the solution from EHC denoted by \textbf{EHC-CGC}.
%
The proposed correlation cluster fusion algorithm with EHC-based and watershed-based proposals
 and MC-I and CGC as subproblem solvers denoted by 
\textbf{fusionCC-HC-MC},
\textbf{-HC-CGC},
\textbf{-WS-MC}, and 
\textbf{-WS-CGC}, respectively.
We set the size of the proposals fix to $10\%$ of the original problem size
and the standard derivation of the random edge noise to $1.5$.

 \begin{figure}[t]
    \centering
    \begin{tikzpicture}
    \begin{semilogxaxis}[  mark size=1pt,
     height= 4cm, 
    restrict y to domain=-80000:1000,
   %   restrict x to domain=10:4000,
    xlabel = {runtime},
    xmin = 10,
    xmax = 4100,
    ymax = 0,
    ymin = -80000,
    legend to name = ledgendPositionMC,
    legend columns=4,
    width = 0.9\columnwidth,
    scaled ticks = false
    ] 
    \addplot[color=red,mark=x] table[x=time, y=MC-CCFDB-o*]{data3/knott-3d-450b.data};          \addlegendentry{MCR-old}
    \addplot[color=orange,mark=o] table[x=time, y=MC-CCFDB-n*]{data3/knott-3d-450b.data};          \addlegendentry{MCR-new}
    \addplot[color=blue,mark=x] table[x=time, y=MC-CCIFD-o*]{data3/knott-3d-450b.data};          \addlegendentry{MCI-old}
    \addplot[color=cyan,mark=o] table[x=time, y=MC-CCIFD-n*]{data3/knott-3d-450b.data};          \addlegendentry{MCI-new}

    \end{semilogxaxis}
    \end{tikzpicture}
  \begin{center}
  \hypersetup{linkcolor = black}
  \ref{ledgendPositionMC}
  \hypersetup{linkcolor = red}
  \end{center}
    \caption{Comparison of the Multicut implementation of OpenGM and 
    our modified implementation, which can deal better with large scale problems. 
    However, for large scale problems it still does not scale.}
    \label{fig:at:knott-450b}
  \end{figure}

@@

>>>>

As classical fusion, CCFusion does not provide and lower bound on the objective and 
has no sound stopping condition.
For later we set a maximal number of iterations and maximal number of iteration without improvement.

@@

\subsection{Fast Optimization of CCFusion Moves} \label{sec:fast_optimization}
In general the auxiliary fusion problem~\ref{eq:fusion_move_a} is, as for classical fusion~\cite{Lempitsky-2010}, still NP-hard.
However, many variables have been fixed to be zero and we can reformulate~\ref{eq:fusion_move_a}
into a correlation clustering problem on a coarser graph, where all nodes which are connected
via must-link constraints are merged into single nodes. We call this graph \emph{contracted graph}.
%
\begin{definition}\emph{(Contracted Graph)}
Given a weighted graph $G=V,E,w)$ segmentation of $G$ given by $y\in P(G)$,
we denote by $S: V \to \mathbb{N}$ a mapping from the nodes to the segments.
We define the contraction of graph $G_y=(V_y,E_y, \bar{w})$
by $V_y=\{S(v)|v\in V\}$, $E_y=\{S(u)S(v)|uv\in E\}$, and 
$\forall \bar{u}\bar{v}\in E_y: \bar{w}_{\bar{u}\bar{v}}=\sum_{uv\in E, S(u)=\bar{u},S(v)=\bar{v}}w_{uv}$ 
\end{definition}
%
Any clustering $\bar{y}$ of the contracted graph $G_y=(V_y,E_y)$ can be \emph{back projected} to a clustering $y$ the original graph $G=(V,E)$ by
%
\begin{align}
y_{uv}&=\left\{ 
\begin{array}{ll}
\bar{y}_{S(u)S(v)}& \textrm{ if }S(u) \neq S(v)\\
0                & \textrm{ else}
\end{array}
\right.& \forall uv \in E
\end{align}
Furthermore, the back projection of the optimal segmentation $\bar{y}$ of the contracted graph is a solution of problem~\ref{eq:fusion_move_a}.

@@

As discussed in~\cite{Lempitsky-2010}, proposals
should have two properties: \emph{quality} 
and \emph{diversity}.

A proposal have a high quality if it at least in some regions have a low energy.
For high quality proposals the chance that the inner polytope includes 
better solution (vertex) is larger than for those with low quality.

Diversity between the individual proposals increase the chances
to span diverse internal polytopes, cover with the intersection of
inner polytopes a large part of the original polytope and so  
more likely find the global optimal solution or escape from local minima.

For correlation clustering fusion we add a third
property: \emph{size}.
The size of the contracted graph directly
depends on the number of connected components
of the intersection of the proposal solution and the current best solution.
%
In the extreme case, where each 
node is in a separate connnected component,
the fusion move is equivalent to solving the original problem.
In the other extreme, where the proposal has 
a single connected component, the current best solution will not change.
% 
Therefore the size of the proposals
should be small enough, such that solving
eq.~\ref{eq:fusion_move_a} can
be done fast enough, but  on the other side 
large enough to define a large internal polytope and so a powerful move.
%
To this end we suggest two proposal generators.

@@

\noindent \textbf{Randomized HierarchicalClustering (RHC):}
%\subsubsection{Randomized HierarchicalClustering (RHC)}
%
To generate fast energy aware proposals we can use energy
based hierarchical clustering (EHC) as defined in Sec.~\ref{sec:ehc}.
EHC follows the energy function, therefore the \emph{quality}
of the proposals is high.
To get \emph{diversity} among the different proposal,
we add normal distributed noise  $\mathcal{N}(0, \sigma_{ehc})$
to each edge weight.
To get proposals of the desired \emph{size}, we use a
different stop condition as in EHC, and stop iff a certain number of connected
components is reached.

@@

\noindent \textbf{Randomized Watersheds (RWS):}
%\subsubsection{Randomized Watersheds (RWS)}
Watersheds have become quite popular for
graph segmentation and have a strong connection
to energy minimization~\cite{couprie_2011}.
%
The edge weighted watershed algorithm~\cite{meyer_2013}
with random seeds can be used to get
cheap proposals.
To improve \emph{quality} we do not use $n$ seeds distributed uniformly
over all nodes but use the following.
%
We draw $n/2$ negative edges, and assign to the endpoints a different seed.
Doing so, a random subset of negative edges is forced
to be cut within each proposal.
For additional \emph{diversity}, noise $\mathcal{N} (0, sigma_{ws})$
can be added to the edge weighs.

@@

@@

For \textbf{CGC}~\cite{beier_2014_cvpr} we used the public available implementation\footnote{\label{opengm-cvpr2014}\small{{https://github.com/opengm/opengm/tree/cgc-cvpr2014}}}
of the Beier~\etal and for \textbf{KL}~\cite{Kernighan-1970} the implementation in OpenGM\footlabel{opengm-master}{\url{https://github.com/opengm/opengm}}.
For integer multicuts (\textbf{MC-I}) and relaxed multicuts (\textbf{MC-R})~\cite{kappes_2013_arxiv} we use an modified version of
 the public available implementation in OpenGM\footref{opengm-master} of Kappes~\etal, as described in Sec.~\ref{sec:tunedMC}.
%
%We realize that their implementation has some limitations on large problems and make the following modifications.
%Firstly, we use \emph{index-min-heap}~\cite{Sedgewick-2011} within the shortest path search by the Dykstra algorithm - this gives a speed up for MC-R.
%Secondly, we follow~\cite{kroeger_2012_eccv} and search for shortest paths and add those only if they are non-chordal instead of searching for the shortest non-chordal path
%during the separation procedure. Fig.~\ref{fig:at:knott-450b} show the improvements with our modifications compared to the implementation in OpenGM for the knott-3d-450 dataset.
%This procedure is one magnitude faster, but might cause a few extra outer iterations.
%In~\cite{kroeger_2012_eccv}. this has been used for MC-I only. 
%For MC-R this search procedure is not sufficient and need to be followed by a search for a shortest non-chordal paths.
%
From the field of data-mining we compare to the PIVOT-algorithm~\cite{Ailon-2008} followed by a round of BOEM~\cite{Gionis-2007} denoted by \textbf{PIVOT-BOEM}.
We use the public available code of Micha Elsner\footnote{\url{http://www.ling.ohio-state.edu/~melsner/resources/correlation-readme.html}}.
While this implementation uses full adjacency matrixes it does not scale as well as the algorithm can and therefor can not be applied on all datasets.
We also run classical fusion moves~\cite{Lempitsky-2010} (\textbf{Fusion}) and select distinct labels for the two candidate segmentations.

@@

We compare this methods to the following methods suggested in the present paper.
%
Energy Based Based Hierarchical Clustering (\textbf{EHC})
as described in sec.~\ref{sec:ehc}.
%
CGC warm started with the solution from EHC denoted by \textbf{EHC-CGC}.
%
The proposed correlation cluster fusion algorithm with EHC-based and watershed-based proposals
 and MC-I and CGC as subproblem solvers denoted by 
\textbf{CCFusion-HC-MC},
\textbf{-HC-CGC},
\textbf{-WS-MC}, and 
\textbf{-WS-CGC}, respectively.
We set the size of the proposals fix to $10\%$ of the original problem size
and the standard derivation of the random edge noise to $1.5$.

@@

\subsection{Tuned Multicut Algorithm}\label{sec:tunedMC} 
 \begin{figure}[t]
    \centering
    \begin{tikzpicture}
    \begin{semilogxaxis}[  mark size=1pt,
     height= 4cm, 
    restrict y to domain=-80000:1000,
   %   restrict x to domain=10:4000,
    xlabel = {runtime},
    xmin = 10,
    xmax = 4100,
    ymax = 0,
    ymin = -80000,
    legend to name = ledgendPositionMC,
    legend columns=4,
    width = 0.9\columnwidth,
    scaled ticks = false
    ] 
    \addplot[color=red,mark=x] table[x=time, y=MC-CCFDB-o*]{data3/knott-3d-450b.data};          \addlegendentry{MCR-old}
    \addplot[color=orange,mark=o] table[x=time, y=MC-CCFDB-n*]{data3/knott-3d-450b.data};          \addlegendentry{MCR-new}
    \addplot[color=blue,mark=x] table[x=time, y=MC-CCIFD-o*]{data3/knott-3d-450b.data};          \addlegendentry{MCI-old}
    \addplot[color=cyan,mark=o] table[x=time, y=MC-CCIFD-n*]{data3/knott-3d-450b.data};          \addlegendentry{MCI-new}

    \end{semilogxaxis}
    \end{tikzpicture}
  \begin{center}
  \hypersetup{linkcolor = black}
  \ref{ledgendPositionMC}
  \hypersetup{linkcolor = red}
  \end{center}
    \caption{Comparison of the Multicut implementation of OpenGM and 
    our modified implementation, which can deal better with large scale problems. 
    However, for large scale problems it still does not scale.}
    \label{fig:at:knott-450b}
  \end{figure}
When using the  public available implementation in OpenGM\footref{opengm-master}
of Kappes~\etal~\cite{kappes_2013_arxiv}, we realize that their implementation has some limitations on large problems.
This results in a very slow separation and make the following modifications.
Firstly, we use \emph{index-min-heap}~\cite{Sedgewick-2011} within the shortest path search by the Dykstra algorithm - this gives a speed up for MC-R.
Secondly, we follow~\cite{kroeger_2012_eccv} and search for shortest paths and add those only if they are non-chordal instead of searching for the shortest non-chordal path
during the separation procedure. Fig.~\ref{fig:at:knott-450b} show the improvements with our modifications compared to the implementation in OpenGM for the knott-3d-450 dataset.
This procedure is one magnitude faster, but might cause a few extra outer iterations.
In~\cite{kroeger_2012_eccv}. this has been used for MC-I only. 
For MC-R this search procedure is not sufficient and need to be followed by a search for a shortest non-chordal paths.
